HICF1 - factor modelling - TREES - V2
========================================================

Dr. Susanne Weller 
27/05/2014

Import merged dataframe - v2

```{r}
setwd("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v3")
load("genclinv3.Rda")
library(plyr)

#reorder collumns for convenience
genclinv3<-genclinv3[c(1, 4, 8, 2, 6, 5, 50, 3, 7, 9:11, 13, 15:50, 53:57 )]
genclinv3$X11q_monoallelic_mut <- NULL
genclinv3$Binet.x.1 <- NULL
genclinv3 <-rename(genclinv3, c("X11q_monoallelic_del.y"="X11q_mono_del", "X11q_biallelic.y"="X11q_biallelic"))
names(genclinv3)[7]<- "Binet"
                                
#convert integers to factors

genclinv3$vhmut <- as.factor(genclinv3$vhmut)
levels(genclinv3$MRD) <- c("-MRD", "+MRD")
genclinv3$cd38 <- as.factor(genclinv3$cd38)
genclinv3[,13:47] <- lapply(genclinv3[,13:47], as.factor)

#combine genetic factors
genclinv3$SAMHD1 <- factor(ifelse(genclinv3$SAMHD1_1mutationONLY == "1"|genclinv3$SAMHD1_biallelic_all =="1", "1", "0"))
genclinv3$TP53 <- factor(ifelse(genclinv3$TP53_mutationONLY == "1"|genclinv3$TP53biallelic =="1", "1", "0"))
genclinv3 <- genclinv3[c(1:47, 53:54, 48:52)]

```

SIMPLE CLASSIFICATION TREES - with blood data added
===================================================

```{r}
treegenclin <- genclinv3[c(6:54)]

library(rpart)
library(rpart.plot)
source("lastfunction.R")

pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v3/figures_trees_03062014/01_MissClassErr_alldata.pdf")
par(mfrow=c(1, 1))
for (cp in c(0.01, 0.009, 0.008, 0.006, 0.004, 0.002, 0.001)){
  source("lastfunction.R")
  nobuckets <- NULL
  MissClassErr <- NULL
  CorrectClass <- NULL
  rootnoderr <- NULL
  relErr <- NULL
  for (i in 1:15) {
    # built the tree
    rp.treegenclin <- rpart(treegenclin$MRD~., method="class", data=treegenclin, control=rpart.control(minbucket=i, xval=10, cp=cp))
    # get the number of buckets
    nobuckets[i] <- i
    # get the root node error
    rootnoderr[i] <- rp.treegenclin$parms$prior[2]
    # Get the rel error for each tree
    rpsum <-printcp(rp.treegenclin)
    relErr[i] <-last(rpsum)
    # Calculate missclassification error
    MissClassErr[i] <-rootnoderr[i]*relErr[i]
    # Calculate correctly classified patients
    CorrectClass[i] <- 209 - MissClassErr[i]*209
    }
MissClassErr
if (cp==0.01){
plot(nobuckets, MissClassErr, pch=0, type="o", ylim=c(0, 0.4), main="All data without age, treatment, freqvhmut", xlab="No. of patients in end group\n(more patients=less complex model)", ylab="Missclassification error=fraction of patients that will be put in wrong group ")
legend(12, 0.2, c("0.01", "0.009", "0.008", "0.006", "0.004", "0.002", "0.001"), cex=0.8, col=c("black","green", "blue", "orange", "purple", "darkslategray4", "red"), pch=0:6, title="complexity:\n lower = more complex")
}
if (cp==0.009){
lines(x=nobuckets+0.05, MissClassErr, type="o", pch=1, col="green")
}
if (cp==0.008){
lines(x=nobuckets+0.1, MissClassErr, type="o", pch=2, col="blue")
}
if (cp==0.006){
lines(x=nobuckets+0.15, MissClassErr, type="o", pch=3, col="orange")
}  
if (cp==0.004){
lines(x=nobuckets+0.20, MissClassErr, type="o", pch=4, col="purple")
}
if (cp==0.002){
lines(x=nobuckets+0.25, MissClassErr, type="o", pch=5, col="darkslategray4")
}
if (cp==0.001){
lines(x=nobuckets+0.30,MissClassErr, type="o", pch=6, col="red")
}
}
dev.off()


pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v3/figures_trees_03062014/02_tree_alldata_cp001_buk1.pdf")
#plot best tree with these variables
rp.treegenclinbuk3cp008 <- rpart(treegenclin$MRD~., method="class", data=treegenclin, control=rpart.control(minbucket=1, xval=10, cp=0.001))
prp(rp.treegenclinbuk3cp008, extra=8, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="All data without age, treatment, freqvhmut \n cp=0.001, endgroup=1, MissClassErr~0.01% ")
dev.off()

pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v3/figures_trees_03062014/03_tree_alldata_cp01_buk5.pdf")
rp.treegenclinbuk5cp01 <- rpart(treegenclin$MRD~., method="class", data=treegenclin, control=rpart.control(minbucket=5, xval=10, cp=0.01))
prp(rp.treegenclinbuk5cp01, extra=2, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="All data without age, treatment, freqvhmut \n cp=0.01, endgroup=5, MissClassErr~15% ", )
dev.off()


#rp.treegenclinbuk2cp001 <- rpart(treegenclin$MRD~., method="class", data=treegenclin, control=rpart.control(minbucket=2, xval=10, cp=0.001))
#pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/04_tree_alldata_cp001_buk2.pdf")
# prp(rp.treegenclinbuk3cp008, extra=2, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="All data without age, treatment, freqvhmut \n cp=0.008, endgroup=3, MissClassErr~21% ", )
# dev.off()
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/05_tree_alldata_cp001_buk2.pdf")
# prp(rp.treegenclinbuk2cp001, extra=8, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="All data without age, treatment, freqvhmut \n cp=0.001, endgroup=2, MissClassErr~12% ", )
# dev.off()
```
Checking frequency distributions of blood data:
===============================================
```{r}

pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v3/figures_trees_03062014/04_hist_haem.pdf")
hist(genclinv3$haem, breaks=24)
abline(v=11,col=3,lty=1)
abline(v=9.1,col=4,lty=1)
dev.off()
pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v3/figures_trees_03062014/05_hist_plat.pdf")
hist(genclinv3$plat, breaks=24)
abline(v=148,col=4,lty=1)
dev.off()
pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v3/figures_trees_03062014/06_hist_WBC.pdf")
hist(genclinv3$WBC, breaks=20)
abline(v=289,col=4,lty=1)
abline(v=76,col=4,lty=1)
dev.off()
pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v3/figures_trees_03062014/07_hist_ANC.pdf")
hist(genclinv3$ANC_Neutro, breaks=15)
abline(v=52,col=4,lty=1)
dev.off()
pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v3/figures_trees_03062014/08_hist_ALC.pdf")
hist(genclinv3$ALC_Lympho, breaks=20)
abline(v=20,col=4,lty=1)
dev.off()
```
Principal component analysis for blood data:
============================================

Objective:
Check if all blood variables point in the same direction, e.g. could be reduced to one variable

```{r}

require(FactoMineR) 
# PCA with function PCA

pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v3/figures_trees_03062014/09_pca_blood.pdf")
#Classify blood variables
blood_pca <- PCA(genclinv3[50:54], scale.unit=TRUE, ncp=5, graph=T)
dev.off()

#scale all the features,  ncp: number of dimensions kept in the results (by default 5)
#blood_load <- dimdesc(blood_pca)
#patient_load
#blood_importance_c1 <- as.data.frame(blood_load$Dim.1$quanti)


km.out=kmeans(genclinv3[50:54],2,nstart=10, na.omit=TRUE)
km.out


# William 2011 (Clin. Oncol.)
# ==========================
# - 17q del (FISH)
# - 11q del (FISH)
# - vhmut
# - gender
# - tri12
# - cd38
# - 12q del (FISH)
# 
# ```{r}
# treewilliam <- genclinv2[c(6, 8, 9, 10, 19, 46, 35)]
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/06_MissClassErr_william.pdf")
# par(mfrow=c(1, 1))
# 
# for (cp in c(0.01, 0.009, 0.008, 0.006, 0.004, 0.002, 0.001)){
#   source("lastfunction.R")
#   nobuckets <- NULL
#   MissClassErr <- NULL
#   CorrectClass <- NULL
#   rootnoderr <- NULL
#   relErr <- NULL
#   for (i in 1:20) {
#     # built the tree
#     rp.treewilliam <- rpart(treewilliam$MRD~., method="class", data=treewilliam, control=rpart.control(minbucket=i, xval=10, cp=cp))
#     # get the number of buckets
#     nobuckets[i] <- i
#     # get the root node error
#     rootnoderr[i] <- rp.treewilliam$parms$prior[2]
#     # Get the rel error for each tree
#     rpsum <-printcp(rp.treewilliam)
#     relErr[i] <-last(rpsum)
#     # Calculate missclassification error
#     MissClassErr[i] <-rootnoderr[i]*relErr[i]
#     # Calculate correctly classified patients
#     CorrectClass[i] <- 209 - MissClassErr[i]*209
#     }
# if (cp==0.01){
# plot(nobuckets, MissClassErr, pch=0, type="o", ylim=c(0, 0.4), main="Variables used in William 2011", xlab="No. of patients in end group\n(more patients=less complex model)", ylab="Missclassification error=fraction of patients that will be put in wrong group ")
# legend(12, 0.2, c("0.01", "0.009", "0.008", "0.006", "0.004", "0.002", "0.001"), cex=0.8, col=c("black","green", "blue", "orange", "purple", "darkslategray4", "red"), pch=0:7, title="complexity:\n lower = more complex")
# }
# if (cp==0.009){
# lines(x=nobuckets+0.05, MissClassErr, type="o", pch=1, col="green")
# }
# if (cp==0.008){
# lines(x=nobuckets+0.1, MissClassErr, type="o", pch=2, col="blue")
# }
# if (cp==0.006){
# lines(x=nobuckets+0.15, MissClassErr, type="o", pch=3, col="orange")
# }  
# if (cp==0.004){
# lines(x=nobuckets+0.20, MissClassErr, type="o", pch=4, col="purple")
# }
# if (cp==0.002){
# lines(x=nobuckets+0.25, MissClassErr, type="o", pch=5, col="darkslategray4")
# }
# if (cp==0.001){
# lines(x=nobuckets+0.30,MissClassErr, type="o", pch=6, col="red")
# }
# }
# dev.off()
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/07_tree_william_cp001_buk4.pdf")
# #plot best tree with these variables
# rp.treewilliam <- rpart(treewilliam$MRD~., method="class", data=treewilliam, control=rpart.control(minbucket=4, xval=10, cp=0.01))
# prp(rp.treewilliam, extra=8, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Variables used in William 2011 \n cp=0.001, endgroup=4, MissClassErr~34% ")
# dev.off()
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/08_tree_william_cp001_buk4.pdf")
# prp(rp.treewilliam, extra=2, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Variables used in William 2011 \n cp=0.001, endgroup=4, MissClassErr~34% ")
# dev.off()
# ```
# Rossi 2014 (Blood)
# ==========================
# - TP53 disruption *(TP53_mutationONLY, TP53biallelic, TP53)*
# - BIRC3 disruption (45)
# - SF3B1 mutation
# - NOTCH1 mutation
# - 11q del (46)
# ```{r}
# treerossi <- genclinv2[c(9, 15, 16, 17, 18, 46, 49, 45)]
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/09_MissClassErr_rossi.pdf")
# par(mfrow=c(1, 1))
# for (cp in c(0.01, 0.008, 0.006, 0.004, 0.002, 0.001)){
#   missclassbybucket <- NULL
#   nobuckets <- NULL
#   MissClassErr <- NULL
#   CorrectClass <- NULL
#   rootnoderr <- NULL
#   for (i in 1:20) {
#     # built the tree
#     rp.treerossi <- rpart(treerossi$MRD~., method="class", data=treerossi, control=rpart.control(minbucket=i, xval=10, cp=cp))
#     # get the number of buckets
#     nobuckets[i] <- i
#     # get the root node error
#     rootnoderr[i] <- rp.treerossi$parms$prior[2]
#     # Get the rel error for each tree
#     rpsum <-printcp(rp.treerossi)
#     relErr <-last(rpsum)
#     # Calculate missclassification error
#     MissClassErr[i] <-rootnoderr[i]*relErr
#     # Calculate correctly classified patients
#     CorrectClass[i] <- 209 - MissClassErr[i]*209
#     }
# if (cp==0.01){
# plot(nobuckets, MissClassErr, pch=0, type="o", ylim=c(0, 0.6), main="Variables used in Rossi 2014", xlab="No. of patients in end group\n(more patients=less complex model)", ylab="Missclassification error=fraction of patients that will be put in wrong group ")
# legend(12, 0.3, c("0.01", "0.008", "0.006", "0.004", "0.002", "0.001"), cex=0.8,col=c("black","green", "blue", "orange", "purple", "red"), pch=0:5, title="complexity \n low-> more complex")
# }
# if (cp==0.008){
# lines(MissClassErr, type="o", pch=1, col="green")
# }
# if (cp==0.006){
# lines(MissClassErr, type="o", pch=2, col="blue")
# }
# if (cp==0.004){
# lines(MissClassErr, type="o", pch=3, col="orange")
# }  
# if (cp==0.002){
# lines(MissClassErr, type="o", pch=4, col="purple")
# }
# if (cp==0.001){
# lines(MissClassErr, type="o", pch=5, col="red")
# }
# }
# dev.off()
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/10_tree_rossi_cp001_buk2.pdf")
# #plot best tree with these variables
# rp.treerossi <- rpart(treerossi$MRD~., method="class", data=treerossi, control=rpart.control(minbucket=2, xval=10, cp=0.001))
# prp(rp.treerossi, extra=2, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Variables used in Rossi 2014 \n cp=0.001, minbucket= 2, MissClassErr~38%")
# dev.off()
# 
# ```
# Rosenquist 2014 (Leukemia)
# ==========================
# - TP53 *(TP53_mutationONLY, TP53biallelic, TP53)*
# - BIRC3
# - SF3B1
# - NOTCH1
# - del 17p (35)*(X17q11q12)*
# - del 11q (46)
# - trisomy 12 (19)
# - del 13q (39:41) *(X13q_hom. X13q_het, X13q_redefinedMRD_loss_4)*
# - MYD88 (24)
# - CNA (12)
# ```{r}
# treerosenq <- genclinv2[c(9, 15, 16, 17, 18, 19, 45, 24, 35, 39:41, 46, 49, 12)]
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/11_MissClassErr_rosenquist.pdf")
# par(mfrow=c(1, 1))
# for (cp in c(0.01, 0.008, 0.006, 0.004, 0.002, 0.001)){
#   missclassbybucket <- NULL
#   nobuckets <- NULL
#   MissClassErr <- NULL
#   CorrectClass <- NULL
#   rootnoderr <- NULL
#   for (i in 1:20) {
#     # built the tree
#     rp.treerosenq <- rpart(treerosenq$MRD~., method="class", data=treerosenq, control=rpart.control(minbucket=i, xval=10, cp=cp))
#     # get the number of buckets
#     nobuckets[i] <- i
#     # get the root node error
#     rootnoderr[i] <- rp.treerosenq$parms$prior[2]
#     # Get the rel error for each tree
#     rpsum <-printcp(rp.treerosenq)
#     relErr <-last(rpsum)
#     # Calculate missclassification error
#     MissClassErr[i] <-rootnoderr[i]*relErr
#     # Calculate correctly classified patients
#     CorrectClass[i] <- 209 - MissClassErr[i]*209
#     }
# if (cp==0.01){
# plot(nobuckets, MissClassErr, pch=0, type="o", ylim=c(0, 0.5), main="Variables used in Rosenquist 2014", xlab="No. of patients in end group\n(more patients=less complex model)", ylab="Missclassification error=fraction of patients that will be put in wrong group ")
# legend(12, 0.3, c("0.01", "0.008", "0.006", "0.004", "0.002", "0.001"), cex=0.8,col=c("black","green", "blue", "orange", "purple", "red"), pch=0:5, title="complexity \n low-> more complex")
# }
# if (cp==0.008){
# lines(MissClassErr, type="o", pch=1, col="green")
# }
# if (cp==0.006){
# lines(MissClassErr, type="o", pch=2, col="blue")
# }
# if (cp==0.004){
# lines(MissClassErr, type="o", pch=3, col="orange")
# }  
# if (cp==0.002){
# lines(MissClassErr, type="o", pch=4, col="purple")
# }
# if (cp==0.001){
# lines(MissClassErr, type="o", pch=5, col="red")
# }
# }
# dev.off()
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/12_tree_rosenq_cp001_buk1.pdf")
# #plot best tree with these variables
# rp.treerosenq <- rpart(treerosenq$MRD~., method="class", data=treerosenq, control=rpart.control(minbucket=1, xval=10, cp=0.001))
# prp(rp.treerosenq, extra=2, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Variables used in Rosenquist 2014 \n cp=0.001, minbucket= 1, MissClassErr~25%")
# dev.off()
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/13_tree_rosenq_cp001_buk3.pdf")
# rp.treerosenq <- rpart(treerosenq$MRD~., method="class", data=treerosenq, control=rpart.control(minbucket=3, xval=10, cp=0.001))
# prp(rp.treerosenq, extra=2, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Variables used in Rosenquist 2014 \n cp=0.001, minbucket= 3, MissClassErr~25%")
# dev.off()
# 
# ```
# Health Economic Plan (HEP)
# ==========================
# - TP53 *(TP53_mutationONLY, TP53biallelic, TP53)*
# - Isolated del_13q (39:41)  *(X13q_hom. X13q_het, X13q_redefinedMRD_loss_4)*
# - Isolated trisomy 12 (19)
# - NOTCH1 (18)
# - del_11q (46)
# - SF3B1 (17)
# - BIRC3 del *(BIRC3_mutation)*
# - subclones (11)
# - SAMHD1 (13, 14)
# - MED12
# 
# - TP53 
# - BIRC3
# - SF3B1 (17)
# - NOTCH1 (18)
# - del 17p (35)*(X17q11q12)*
# - del 11q (46)
# - trisomy 12 (19)
# - del 13q 
# - MYD88 (24)
# 
# ```{r}
# treehep <- genclinv2[c(9, 15, 16, 49, 39:41, 19, 18, 46, 17, 11, 13, 14, 25)]
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/14_MissClassErr_hep.pdf")
# par(mfrow=c(1, 1))
# for (cp in c(0.01, 0.008, 0.006, 0.004, 0.002, 0.001)){
#   missclassbybucket <- NULL
#   nobuckets <- NULL
#   MissClassErr <- NULL
#   CorrectClass <- NULL
#   rootnoderr <- NULL
#   for (i in 1:20) {
#     # built the tree
#     rp.treehep <- rpart(treehep$MRD~., method="class", data=treehep, control=rpart.control(minbucket=i, xval=10, cp=cp))
#     # get the number of buckets
#     nobuckets[i] <- i
#     # get the root node error
#     rootnoderr[i] <- rp.treehep$parms$prior[2]
#     # Get the rel error for each tree
#     rpsum <-printcp(rp.treehep)
#     relErr <-last(rpsum)
#     # Calculate missclassification error
#     MissClassErr[i] <-rootnoderr[i]*relErr
#     # Calculate correctly classified patients
#     CorrectClass[i] <- 209 - MissClassErr[i]*209
#     }
# if (cp==0.01){
# plot(nobuckets, MissClassErr, pch=0, type="o", ylim=c(0, 0.5), main="Variables proposed for Health Economic Plan", xlab="No. of patients in end group\n(more patients=less complex model)", ylab="Missclassification error=fraction of patients that will be put in wrong group ")
# legend(12, 0.3, c("0.01", "0.008", "0.006", "0.004", "0.002", "0.001"), cex=0.8,col=c("black","green", "blue", "orange", "purple", "red"), pch=0:5, title="complexity \n low-> more complex")
# }
# if (cp==0.008){
# lines(MissClassErr, type="o", pch=1, col="green")
# }
# if (cp==0.006){
# lines(MissClassErr, type="o", pch=2, col="blue")
# }
# if (cp==0.004){
# lines(MissClassErr, type="o", pch=3, col="orange")
# }  
# if (cp==0.002){
# lines(MissClassErr, type="o", pch=4, col="purple")
# }
# if (cp==0.001){
# lines(MissClassErr, type="o", pch=5, col="red")
# }
# }
# dev.off()
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/15_tree_hep_cp001_buk1.pdf")
# #plot best tree with these variables
# rp.treehep <- rpart(treehep$MRD~., method="class", data=treehep, control=rpart.control(minbucket=1, xval=10, cp=0.001))
# prp(rp.treehep, extra=8, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Variables proposed for Health Economics Plan \n cp=0.001, endgroup=1, MissClassErr~27%")
# dev.off()
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/16_tree_hep_cp001_buk1.pdf")
# prp(rp.treehep, extra=2, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Variables proposed for Health Economics Plan \n cp=0.001, endgroup=1, MissClassErr~27%")
# dev.off()
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/17_tree_hep_cp001_buk4.pdf")
# rp.treehepb <- rpart(treehep$MRD~., method="class", data=treehep, control=rpart.control(minbucket=4, xval=10, cp=0.001))
# prp(rp.treehepb, extra=2, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Variables proposed for Health Economics Plan \n cp=0.001, endgroup=4, MissClassErr~31%")
# dev.off()
# ```
# Only genetic data
# =================
# ```{r}
# treegenetic <- genclinv2[c(9, 11:49)]
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/18_MissClassErr_genetics.pdf")
# 
# par(mfrow=c(1, 1))
# for (cp in c(0.01, 0.009, 0.008, 0.007, 0.006, 0.004, 0.002, 0.001)){
#   nobuckets <- NULL
#   MissClassErr <- NULL
#   CorrectClass <- NULL
#   rootnoderr <- NULL
#   relErr <- NULL
#   print(cp)
#   for (i in 1:15) {
#     # built the tree
#     rp.treegenetic <- rpart(treegenetic$MRD~., method="class", data=treegenetic, control=rpart.control(minbucket=i, xval=10, cp=cp))
#     # get the number of buckets
#     nobuckets[i] <- i
#     # get the root node error
#     rootnoderr[i] <- rp.treegenetic$parms$prior[2]
#     # Get the rel error for each tree
#     rpsum <-printcp(rp.treegenetic)
#     relErr[i] <-last(rpsum)
#     # Calculate missclassification error
#     MissClassErr[i] <-rootnoderr[i]*relErr[i]
#     # Calculate correctly classified patients
#     CorrectClass[i] <- 209 - MissClassErr[i]*209
#     }
# if (cp==0.01){
# plot(nobuckets, MissClassErr, pch=0, type="o", ylim=c(0, 0.4), main="Only our genetic findings", xlab="No. of patients in end group\n(more patients=less complex model)", ylab="Missclassification error=fraction of patients that will be put in wrong group")
# legend(12, 0.25, c("0.01", "0.009", "0.008", "0.007", "0.006", "0.004", "0.002", "0.001"), cex=0.8, col=c("black", "green", "darkslategray4", "darkgoldenrod", "blue", "orange", "purple", "red"), pch=0:7, title="complexity \n low-> more complex")
# }
# if (cp==0.009){
# lines(MissClassErr, type="o", pch=1, col="green")
# }
# if (cp==0.008){
# lines(MissClassErr, type="o", pch=2, col="darkslategray4")
# }
# if (cp==0.007){
# lines(MissClassErr, type="o", pch=3, col="darkgoldenrod")
# }
# if (cp==0.006){
# lines(MissClassErr, type="o", pch=4, col="blue")
# }
# if (cp==0.004){
# lines(MissClassErr, type="o", pch=5, col="orange")
# }  
# if (cp==0.002){
# lines(MissClassErr, type="o", pch=6, col="purple")
# }
# if (cp==0.001){
# lines(MissClassErr, type="o", pch=7, col="red")
# }
# }
# dev.off()
# 
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/19_tree_genetic_cp001_buk1.pdf")
# #plot best tree with these variables
# rp.treegenetic <- rpart(treegenetic$MRD~., method="class", data=treegenetic, control=rpart.control(minbucket=1, xval=10, cp=0.001))
# prp(rp.treegenetic, extra=2, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Only our genetic findings \n cp=0.001, endgroup=1, MissClassErr~11%")
# dev.off()
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/20_tree_genetic_cp001_buk3.pdf")
# rp.treegeneticb <- rpart(treegenetic$MRD~., method="class", data=treegenetic, control=rpart.control(minbucket=3, xval=10, cp=0.001))
# prp(rp.treegeneticb, extra=2, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Only our genetic findings \n cp=0.001, endgroup=5, MissClassErr~22%")
# dev.off()
# 
# pdf("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2/figures_trees_29052014/21_tree_genetic_cp001_buk8.pdf")
# rp.treegeneticc <- rpart(treegenetic$MRD~., method="class", data=treegenetic, control=rpart.control(minbucket=3, xval=10, cp=0.001))
# prp(rp.treegeneticc, extra=2, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Only our genetic findings \n cp=0.001, endgroup=8, MissClassErr~25%")
# dev.off()

