HICF1 CLinical data - descriptive overview
============================================
```{r}
load("/home/suska/work/HICF1/HICF1_sub1/trunk/clinicaldata.Rda")
library(gplots)
library(ggplot2)
```
Is there an age difference between males and females?
```{r}
#using gplots package here
boxplot2(age_ar ~ gender, data=clinicaldata, main="Gender ~ age", col=c("blue", "red"), xlab="gender", ylab="age", top=TRUE, varwidth="T")
#same with ggplot2
g <- ggplot(clinicaldata, aes(gender, age_ar))
give.n <- function(x){
   return(c(y = mean(x), label = length(x)))
}
g + geom_boxplot(col=c("darkblue", "darkred")) + stat_summary(fun.data = give.n, geom = "text")
# independent 2-group t-test with unequal variance (Welsh mod)
t.test(age_ar~gender, data=clinicaldata) # where y is numeric and x is a binary factor
```
Is there an age difference between treatments?
```{r}
t <- ggplot(clinicaldata, aes(treatment, age_ar))
give.n <- function(x){
   return(c(y = mean(x), label = length(x)))
}
t + geom_boxplot(col=c("purple", "darkgreen", "yellow"))+
  stat_summary(fun.data = give.n, geom = "text")+
  ggtitle("Treatment ~ age")+
  ylab("Age (years)")
# independent multi-group Welsh ANOVA
oneway.test(age_ar ~ treatment, data=clinicaldata) # where y is numeric and x is a binary factor
```
Is there an age difference between MRD outcome?
```{r}
clinicaldata_mrd <-na.omit(clinicaldata[,c("mrd","age_ar")])
m <- ggplot(clinicaldata_mrd, aes(mrd, age_ar))
give.n <- function(x){
   return(c(y = mean(x), label = length(x)))
}
m + geom_boxplot(col=c("black", "purple"))+
  stat_summary(fun.data = give.n, geom = "text")+
  ggtitle("MRD ~ age")+
  ylab("Age (years)")
# independent multi-group Welsh ANOVA
t.test(age_ar~mrd, data=clinicaldata)
```
Is there an age difference between response?
- Note here that there are only few datapoints for SD, PD and CR_CRi seem to be older.
- Unfortunately, those might be also very interesting cases...
```{r}
clinicaldata_response <-na.omit(clinicaldata[,c("response","age_ar")])
r <- ggplot(clinicaldata_response, aes(response, age_ar))
give.n <- function(x){
   return(c(y = mean(x), label = length(x)))
}
r + geom_violin(fill = "grey80", colour = "#3366FF")+
  stat_summary(fun.data = give.n, geom = "text")+
  ggtitle("Response ~ age")+
  ylab("Age (years)")
# pairwise Wilcoxon Test with plain Bonferroni
pairwise.wilcox.test(clinicaldata$age, clinicaldata$response, p.adjust.method="bonferroni", paired = FALSE)
```
Frequency analysis of chromosome lesions, mrd and response
```{r}
sum_vhmut_mrd <- table(clinicaldata$vhmut, clinicaldata$mrd)
#colnames(sum_vhmut_mrd) <- c("vhmut", "mrd", "freq") 
chi_vhmut_mrd <- chisq.test(sum_vhmut_mrd)
```
- again here, there are only few datapoints for "biclonal" and "equivocal" vhmut...
-> are these the same as the SD, PD and CR_CRi in response?

```{r}
sub_vhmut_mrd <- subset(clinicaldata, response == "SD" | response == "PD"| response == "CR_CRi" )
```

Look up: Ordered factors in R, necessary?

Use vcr package (see tutorial pdf)

Test variables for independence 
-> Categorical variables with CHiSq or XXX
-> Continuous with correlations 

