HICF1 - factor modelling - TREES - V2
========================================================

Dr. Susanne Weller 
27/05/2014

Import merged dataframe - v2

```{r}
setwd("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees/HICF1_trees_v2")
load("genclinv2.Rda")
library(plyr)

#reorder collumns for convenience
genclinv2<-genclinv2[c(1, 4, 8, 2, 6, 5, 50, 3, 7, 9:11, 13, 15:49 )]
genclinv2$X11q_monoallelic_mut <- NULL
genclinv2 <-rename(genclinv2, c("X11q_monoallelic_del.y"="X11q_mono_del", "X11q_biallelic.y"="X11q_biallelic"))

                                
#convert integers to factors

genclinv2$vhmut <- as.factor(genclinv2$vhmut)
levels(genclinv2$MRD) <- c("-MRD", "+MRD")
genclinv2$cd38 <- as.factor(genclinv2$cd38)
genclinv2[,13:48] <- lapply(genclinv2[,13:48], as.factor)

#combine genetic factors
genclinv2$SAMHD1 <- factor(ifelse(genclinv2$SAMHD1_1mutationONLY == "1"|genclinv2$SAMHD1_biallelic_all =="1", "1", "0"))
genclinv2$TP53 <- factor(ifelse(genclinv2$TP53_mutationONLY == "1"|genclinv2$TP53biallelic =="1", "1", "0"))

```

SIMPLE CLASSIFICATION TREES
==============================

I. Using Rpart
-------------------------------

*as agreed on 20/05/2014*

Not using age, treatment, freqvhmut

```{r}
treegenclin <- genclinv2[c(6:49)]

library(rpart)
library(rpart.plot)
#rp.treegenclin <- rpart(treegenclin$MRD~., method="class", data=treegenclin, control=rpart.control(minbucket=3, xval=10, cp=0.001))
#par(mfrow=c(1, 1))
#prp(rp.treegenclin, extra=8)
##prp(rp.treegenclin, extra=8)
#prp(rp.treegenclin, extra=8, fallen.leaves=TRUE)

#par(mfrow=c(1, 1))
#prp(rp.treegenclin, extra=8, uniform=TRUE, branch=1, left=FALSE, varlen=0)
#prp(rp.treegenclin, extra=2, uniform=FALSE, branch=1, left=FALSE, varlen=0)


#library(rattle)
#par(mfrow=c(1, 1))
#fancyRpartPlot(rp.treegenclin, uniform=FALSE, branch=1, left=FALSE)

```
Produce different models to find optimal number of patient in buckets and optimal cp
====================================================================================
-> The model already contains 10x CV!

Missclassification error is calculated as "Root node error" x "rel. model error"

(1) produce a number of trees iterating over the number of patients in a bucket
(2) Calculate Missclassification Error rate for those trees
```{r}
library(rpart)
library(rpart.plot)
source("lastfunction.R")

par(mfrow=c(1, 1))
for (cp in c(0.01, 0.008, 0.006, 0.004, 0.002, 0.001)){
  missclassbybucket <- NULL
  nobuckets <- NULL
  MissClassErr <- NULL
  CorrectClass <- NULL
  rootnoderr <- NULL
  for (i in 1:15) {
    # built the tree
    rp.treegenclin <- rpart(treegenclin$MRD~., method="class", data=treegenclin, control=rpart.control(minbucket=i, xval=10, cp=cp))
    # get the number of buckets
    nobuckets[i] <- i
    # get the root node error
    rootnoderr[i] <- rp.treegenclin$parms$prior[2]
    # Get the rel error for each tree
    rpsum <-printcp(rp.treegenclin)
    relErr <-last(rpsum)
    # Calculate missclassification error
    MissClassErr[i] <-rootnoderr[i]*relErr
    # Calculate correctly classified patients
    CorrectClass[i] <- 209 - MissClassErr[i]*209
    }
if (cp==0.01){
plot(nobuckets, CorrectClass, pch=0, type="o", ylim=c(150, 209), main="All data without age, treatment, freqvhmut")
}
else {
  if (cp==0.008){
  lines(CorrectClass, type="o", pch=1, col="green")
  }
  else {
    if (cp==0.006){
    lines(CorrectClass, type="o", pch=2, col="blue")
    }
    else {
      if (cp==0.004){
      lines(CorrectClass, type="o", pch=3, col="orange")
      }  
      else {
        if (cp==0.002){
        lines(CorrectClass, type="o", pch=4, col="purple")
      }
        else {
          if (cp==0.001){
          lines(CorrectClass, type="o", pch=5, col="red")
          }
           }}}}}
legend(5, 210, c("0.01", "0.008", "0.006", "0.004", "0.002", "0.001"), cex=0.8, 
   col=c("black","green", "blue", "orange", "purple", "red"), pch=0:5)
}
#plot best tree with these variables
rp.treegenclinbuk1cp001 <- rpart(treegenclin$MRD~., method="class", data=treegenclin, control=rpart.control(minbucket=1, xval=10, cp=0.001))
prp(rp.treegenclinbuk1cp001, extra=8, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="All data without age, treatment, freqvhmut \n cp=0.001")
```
William 2011 (Clin. Oncol.)
==========================
- 17q del (FISH)
- 11q del (FISH)
- vhmut
- gender
- tri12
- cd38
- 12q del (FISH)

```{r}
treewilliam <- genclinv2[c(6, 8, 9, 10, 19, 46, 35)]

par(mfrow=c(1, 1))
for (cp in c(0.01, 0.008, 0.006, 0.004, 0.002, 0.001)){
  missclassbybucket <- NULL
  nobuckets <- NULL
  MissClassErr <- NULL
  CorrectClass <- NULL
  rootnoderr <- NULL
  for (i in 1:20) {
    # built the tree
    rp.treewilliam <- rpart(treewilliam$MRD~., method="class", data=treewilliam, control=rpart.control(minbucket=i, xval=10, cp=cp))
    # get the number of buckets
    nobuckets[i] <- i
    # get the root node error
    rootnoderr[i] <- rp.treewilliam$parms$prior[2]
    # Get the rel error for each tree
    rpsum <-printcp(rp.treewilliam)
    relErr <-last(rpsum)
    # Calculate missclassification error
    MissClassErr[i] <-rootnoderr[i]*relErr
    # Calculate correctly classified patients
    CorrectClass[i] <- 209 - MissClassErr[i]*209
    }
if (cp==0.01){
plot(nobuckets, CorrectClass, pch=0, type="o", ylim=c(50, 209), main="Variables used in William 2011")
}
else {
  if (cp==0.008){
  lines(CorrectClass, type="o", pch=1, col="green")
  }
  else {
    if (cp==0.006){
    lines(CorrectClass, type="o", pch=2, col="blue")
    }
    else {
      if (cp==0.004){
      lines(CorrectClass, type="o", pch=3, col="orange")
      }  
      else {
        if (cp==0.002){
        lines(CorrectClass, type="o", pch=4, col="purple")
      }
        else {
          if (cp==0.001){
          lines(CorrectClass, type="o", pch=5, col="red")
          }
           }}}}}
legend(5, 210, c("0.01", "0.008", "0.006", "0.004", "0.002", "0.001"), cex=0.8, 
   col=c("black","green", "blue", "orange", "purple", "red"), pch=0:5)
}
#plot best tree with these variables
rp.treewilliam <- rpart(treewilliam$MRD~., method="class", data=treewilliam, control=rpart.control(minbucket=4, xval=10, cp=0.001))
prp(rp.treewilliam, extra=8, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Variables used in William 2011 \n cp=0.001, minbucket=4")
```
Rossi 2014 (Blood)
==========================
- TP53 disruption *(TP53_mutationONLY, TP53biallelic, TP53)*
- BIRC3 disruption
- SF3B1 mutation
- NOTCH1 mutation
- 11q del (46)
```{r}
treerossi <- genclinv2[c(9, 15, 16, 17, 18, 46)]

par(mfrow=c(1, 1))
for (cp in c(0.01, 0.008, 0.006, 0.004, 0.002, 0.001)){
  missclassbybucket <- NULL
  nobuckets <- NULL
  MissClassErr <- NULL
  CorrectClass <- NULL
  rootnoderr <- NULL
  for (i in 1:20) {
    # built the tree
    rp.treerossi <- rpart(treerossi$MRD~., method="class", data=treerossi, control=rpart.control(minbucket=i, xval=10, cp=cp))
    # get the number of buckets
    nobuckets[i] <- i
    # get the root node error
    rootnoderr[i] <- rp.treerossi$parms$prior[2]
    # Get the rel error for each tree
    rpsum <-printcp(rp.treerossi)
    relErr <-last(rpsum)
    # Calculate missclassification error
    MissClassErr[i] <-rootnoderr[i]*relErr
    # Calculate correctly classified patients
    CorrectClass[i] <- 209 - MissClassErr[i]*209
    }
if (cp==0.01){
plot(nobuckets, CorrectClass, pch=0, type="o", ylim=c(50, 209), main="Variables used in Rossi 2014")
}
else {
  if (cp==0.008){
  lines(CorrectClass, type="o", pch=1, col="green")
  }
  else {
    if (cp==0.006){
    lines(CorrectClass, type="o", pch=2, col="blue")
    }
    else {
      if (cp==0.004){
      lines(CorrectClass, type="o", pch=3, col="orange")
      }  
      else {
        if (cp==0.002){
        lines(CorrectClass, type="o", pch=4, col="purple")
      }
        else {
          if (cp==0.001){
          lines(CorrectClass, type="o", pch=5, col="red")
          }
           }}}}}
legend(5, 210, c("0.01", "0.008", "0.006", "0.004", "0.002", "0.001"), cex=0.8, 
   col=c("black","green", "blue", "orange", "purple", "red"), pch=0:5)
}
#plot best tree with these variables
rp.treerossi <- rpart(treerossi$MRD~., method="class", data=treewilliam, control=rpart.control(minbucket=2, xval=10, cp=0.001))
prp(rp.treewilliam, extra=8, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Variables used in Rossi 2014 \n cp=0.001, minbucket= 2")
```
Rosenquist 2014 (Leukemia)
==========================
- TP53 *(TP53_mutationONLY, TP53biallelic, TP53)*
- BIRC3
- SF3B1
- NOTCH1
- del 17p (35)*(X17q11q12)*
- del 11q (46)
- trisomy 12 (19)
- del 13q (39:41) *(X13q_hom. X13q_het, X13q_redefinedMRD_loss_4)*
- MYD88 (24)
```{r}
treerosenq <- genclinv2[c(9, 15, 16, 17, 18, 19, 45, 24, 35, 39:41, 46, 49)]

par(mfrow=c(1, 1))
for (cp in c(0.01, 0.008, 0.006, 0.004, 0.002, 0.001)){
  missclassbybucket <- NULL
  nobuckets <- NULL
  MissClassErr <- NULL
  CorrectClass <- NULL
  rootnoderr <- NULL
  for (i in 1:20) {
    # built the tree
    rp.treerosenq <- rpart(treerosenq$MRD~., method="class", data=treerosenq, control=rpart.control(minbucket=i, xval=10, cp=cp))
    # get the number of buckets
    nobuckets[i] <- i
    # get the root node error
    rootnoderr[i] <- rp.treerosenq$parms$prior[2]
    # Get the rel error for each tree
    rpsum <-printcp(rp.treerosenq)
    relErr <-last(rpsum)
    # Calculate missclassification error
    MissClassErr[i] <-rootnoderr[i]*relErr
    # Calculate correctly classified patients
    CorrectClass[i] <- 209 - MissClassErr[i]*209
    }
if (cp==0.01){
plot(nobuckets, CorrectClass, pch=0, type="o", ylim=c(50, 209), main="Variables used in Rosenquist 2014")
}
else {
  if (cp==0.008){
  lines(CorrectClass, type="o", pch=1, col="green")
  }
  else {
    if (cp==0.006){
    lines(CorrectClass, type="o", pch=2, col="blue")
    }
    else {
      if (cp==0.004){
      lines(CorrectClass, type="o", pch=3, col="orange")
      }  
      else {
        if (cp==0.002){
        lines(CorrectClass, type="o", pch=4, col="purple")
      }
        else {
          if (cp==0.001){
          lines(CorrectClass, type="o", pch=5, col="red")
          }
           }}}}}
legend(5, 210, c("0.01", "0.008", "0.006", "0.004", "0.002", "0.001"), cex=0.8, 
   col=c("black","green", "blue", "orange", "purple", "red"), pch=0:5)
}
#plot best tree with these variables
rp.treerosenq <- rpart(treerosenq$MRD~., method="class", data=treerosenq, control=rpart.control(minbucket=1, xval=10, cp=0.001))
prp(rp.treerosenq, extra=8, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Variables used in Rosenquist 2014 \n cp=0.001, minbucket= 1")
```
Health Economic Plan (HEP)
==========================
- TP53 *(TP53_mutationONLY, TP53biallelic, TP53)*
- Isolated del_13q (39:41)  *(X13q_hom. X13q_het, X13q_redefinedMRD_loss_4)*
- Isolated trisomy 12 (19)
- NOTCH1 (18)
- del_11q (46)
- SF3B1 (17)
- BIRC3 del *(BIRC3_mutation)*
- subclones (11)
- SAMHD1 (13, 14)
- MED12

- TP53 
- BIRC3
- SF3B1 (17)
- NOTCH1 (18)
- del 17p (35)*(X17q11q12)*
- del 11q (46)
- trisomy 12 (19)
- del 13q 
- MYD88 (24)

```{r}
treehep <- genclinv2[c(9, 15, 16, 49, 39:41, 19, 18, 46, 17, 11, 13, 14, 25)]

par(mfrow=c(1, 1))
for (cp in c(0.01, 0.008, 0.006, 0.004, 0.002, 0.001)){
  missclassbybucket <- NULL
  nobuckets <- NULL
  MissClassErr <- NULL
  CorrectClass <- NULL
  rootnoderr <- NULL
  for (i in 1:20) {
    # built the tree
    rp.treehep <- rpart(treehep$MRD~., method="class", data=treehep, control=rpart.control(minbucket=i, xval=10, cp=cp))
    # get the number of buckets
    nobuckets[i] <- i
    # get the root node error
    rootnoderr[i] <- rp.treehep$parms$prior[2]
    # Get the rel error for each tree
    rpsum <-printcp(rp.treehep)
    relErr <-last(rpsum)
    # Calculate missclassification error
    MissClassErr[i] <-rootnoderr[i]*relErr
    # Calculate correctly classified patients
    CorrectClass[i] <- 209 - MissClassErr[i]*209
    }
if (cp==0.01){
plot(nobuckets, CorrectClass, pch=0, type="o", ylim=c(50, 209), main="Variables proposed for Health Economics Plan")
}
else {
  if (cp==0.008){
  lines(CorrectClass, type="o", pch=1, col="green")
  }
  else {
    if (cp==0.006){
    lines(CorrectClass, type="o", pch=2, col="blue")
    }
    else {
      if (cp==0.004){
      lines(CorrectClass, type="o", pch=3, col="orange")
      }  
      else {
        if (cp==0.002){
        lines(CorrectClass, type="o", pch=4, col="purple")
      }
        else {
          if (cp==0.001){
          lines(CorrectClass, type="o", pch=5, col="red")
          }
           }}}}}
legend(5, 210, c("0.01", "0.008", "0.006", "0.004", "0.002", "0.001"), cex=0.8, 
   col=c("black","green", "blue", "orange", "purple", "red"), pch=0:5)
}
#plot best tree with these variables
rp.treerosenq <- rpart(treerosenq$MRD~., method="class", data=treerosenq, control=rpart.control(minbucket=1, xval=10, cp=0.001))
prp(rp.treerosenq, extra=8, uniform=TRUE, branch=1, left=FALSE, varlen=0, main="Variables proposed for Health Economics Plan \n cp=0.001, minbucket=1")