HICF1 - factor modelling - TREES
========================================================

Dr. Susanne Weller 
16/05/2014

Import merged dataframe

```{r}
setwd("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees")
load("genclin.Rda")

#reorder collumns for convenience
genclin<-genclin[c(1, 7, 2:6, 8:45)]
genclin<-genclin[c(1, 2, 8, 5, 3, 7, 10, 11, 9, 4, 6, 12:45)]


#convert integers to factors
genclin[,9:45] <- lapply(genclin[,9:45], as.factor)
levels(genclin$MRD) <- c("+MRD", "-MRD")


```

USING THE CLASSIC TREE PACKAGE
==============================

I. SIMPLE CLASSIFICATION TREES
-------------------------------

Only genetic data:

```{r}

library(tree)
treegenetic <- genclin[c(2,6,8:45)]
tree.genetic <- tree(MRD~., treegenetic)
summary(tree.genetic)

plot(tree.genetic)
text(tree.genetic, pretty = 0)

```

Using all available data
```{r}
treeall <- genclin[c(2,4:45)]
tree.all <- tree(MRD~., treeall)
summary(tree.all)

plot(tree.all)
text(tree.all, pretty = 0)
```

Using all data except vhmut, freqvhmut, aar
```{r}
wofreq.wovhmut.woaar <- genclin[c(2,4,7:10, 12:45)]
tree.wofreq.wovhmut.woaar <- tree(MRD~., wofreq.wovhmut.woaar)
summary(tree.wofreq.wovhmut.woaar)

plot(tree.wofreq.wovhmut.woaar)
text(tree.wofreq.wovhmut.woaar, pretty = 0)
```

II. SIMPLE CROSS VALIDATION
---------------------------

```{r}

set.seed(5)
train1 <- sample(1:nrow(genclin), 139)
tree.test1 <- genclin[-train1,]
tree.train1 <- genclin[train1,]
MRD.test=treeall$MRD[-train1]

tree.train <- tree(MRD~., treeall, subset=train1)
summary(tree.train)
tree.pred=predict(tree.train, tree.test1, type="class")
table(tree.pred, MRD.test)
```

III. COST COMPLEXITY PRUNING
----------------------------

Using only the training data set

```{r}

set.seed(3)
cv.treetrain <- cv.tree(tree.train, FUN=prune.misclass)
cv.treetrain

par(mfrow=c(1,2))
plot(cv.treetrain$size, cv.treetrain$dev, type="b")
plot(cv.treetrain$k, cv.treetrain$dev, type="b")

prune.train1 <- prune.misclass(tree.train, best=4)
plot(prune.train1)
text(prune.train1, pretty=0)
summary(prune.train1)
```
Note: This pruning results in an improvement of 0.006, possibly due to the fact that the original tree suffered from high variance


Using the whole data set

```{r}
set.seed(5)
cv.tree.all <- cv.tree(tree.all, FUN=prune.misclass)
cv.tree.all

par(mfrow=c(1,2))
plot(cv.tree.all$size, cv.tree.all$dev, type="b")
plot(cv.tree.all$k, cv.tree.all$dev, type="b")

prune.tree.all <- prune.misclass(tree.all, best=4)
plot(prune.tree.all)
text(prune.tree.all, pretty=0)
summary(prune.tree.all)
```

without freqvhmut
```{r}
wofreq <- genclin[c(2,4,5,7:45)]
tree.freq <- tree(MRD~., wofreq)
summary(tree.freq)

plot(tree.freq)
text(tree.freq, pretty = 0)

set.seed(5)
cv.tree.freq <- cv.tree(tree.freq, FUN=prune.misclass)
cv.tree.freq

par(mfrow=c(1,2))
plot(cv.tree.freq$size, cv.tree.freq$dev, type="b")
plot(cv.tree.freq$k, cv.tree.freq$dev, type="b")

prune.tree.freq <- prune.misclass(tree.freq, best=8)
par(mfrow=c(1,1))
plot(prune.tree.freq)
text(prune.tree.freq, pretty=0)
mtext("0.2418%", side=3)
summary(prune.tree.freq)
```

without freqmut, vhmut, age
```{r}
set.seed(5)
cv.tree.wofreq.vhmut.aar <- cv.tree(tree.wofreq.wovhmut.woaar, FUN=prune.misclass)
cv.tree.wofreq.vhmut.aar 

par(mfrow=c(1,2))
plot(cv.tree.wofreq.vhmut.aar$size, cv.tree.wofreq.vhmut.aar$dev, type="b")
plot(cv.tree.wofreq.vhmut.aar$k, cv.tree.wofreq.vhmut.aar$dev, type="b")

prune.tree.wofreq.vhmut.aar <- prune.misclass(tree.wofreq.wovhmut.woaar, best=8)
par(mfrow=c(1,1))
plot(prune.tree.wofreq.vhmut.aar)
text(prune.tree.wofreq.vhmut.aar, pretty=0)
mtext("0.2836%", side=3)
summary(prune.tree.wofreq.vhmut.aar)
```

USING RPART
===========







```
Compare all four boosted models:
```{r}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    for (i in 1:numPlots) {
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
multiplot(boostplot1, boostplot2, boostplot3, boostplot4, cols=2)
