HICF1 - factor modelling - TREES
========================================================

Dr. Susanne Weller 
16/05/2014

Import merged dataframe

```{r}
setwd("~/work/01_HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_trees")
load("genclin.Rda")

#reorder collumns for convenience
genclin<-genclin[c(1, 7, 2:6, 8:45)]
genclin<-genclin[c(1, 2, 8, 5, 3, 7, 10, 11, 9, 4, 6, 12:45)]


#convert integers to factors
genclin[,9:45] <- lapply(genclin[,9:45], as.factor)
levels(genclin$MRD) <- c("-MRD", "+MRD")

#combine genetic factors
genclin$SAMHD1 <- factor(ifelse(genclin$SAMHD1_1mutationONLY == "1"|genclin$SAMHD1_biallelic_all =="1", "1", "0"))
genclin$TP53 <- factor(ifelse(genclin$TP53_mutationONLY == "1"|genclin$TP53biallelic =="1", "1", "0"))

```

SIMPLE CLASSIFICATION TREES
==============================

I. Using Rpart
-------------------------------

*as agreed on 20/05/2014*

Not using age, treatment, frequvhmut

```{r}
treegenclin <- genclin[c(2, 7:14, 19:47)]

library(rpart)
library(rpart.plot)
rp.treegenclin <- rpart(treegenclin$MRD~., method="class", data=treegenclin, control=rpart.control(minbucket=3, xval=10, cp=0.001),)
par(mfrow=c(1, 1))
prp(rp.treegenclin, extra=8)
prp(rp.treegenclin, extra=8)
prp(rp.treegenclin, extra=8, fallen.leaves=TRUE)

par(mfrow=c(1, 1))
prp(rp.treegenclin, extra=8, uniform=FALSE, branch=1, left=FALSE, varlen=0)
prp(rp.treegenclin, extra=2, uniform=FALSE, branch=1, left=FALSE, varlen=0)


library(rattle)
par(mfrow=c(1, 1))
fancyRpartPlot(rp.treegenclin, uniform=FALSE, branch=1, left=FALSE)

```
Cross validate different models to find optimal number of patient in buckets
============================================================================
-> The model already contains 10x CV!

Missclassification error is calculated as "Root node error" x "rel. model error"

(1) produce a number of trees iterating over the number of patients in a bucket
(2) Calculate Missclassification Error rate for those trees
```{r}
library(rpart)
library(rpart.plot)
source("lastfunction.R")

missclassbybucket <- NULL
nobuckets <- NULL
MissClassErr <- NULL
CorrectClass <- NULL
for (i in 1:15) {
# built the tree
rp.treegenclin <- rpart(treegenclin$MRD~., method="class", data=treegenclin, control=rpart.control(minbucket=i, xval=10, cp=0.001))
# get the number of buckets
nobuckets[i] <- i
# get the root node error (always the same, not dependent on number of buckets)
rootnoderr[i] <- rp.treegenclin$parm$prior[2]

# Get the rel error for each tree
rpsum <-printcp(rp.treegenclin)
relErr <-last(rpsum)
# Calculate missclassification error
MissClassErr[i] <-rootnoderr[i]*relErr
# Calculate correctly classified patients
CorrectClass[i] <- 209 - MissClassErr[i]*209
}
 
par(mfrow=c(2, 2))

plot(nobuckets, MissClassErr)
plot(nobuckets, CorrectClass)
rp.treegenclinbuk1 <- rpart(treegenclin$MRD~., method="class", data=treegenclin, control=rpart.control(minbucket=1, xval=10))
prp(rp.treegenclinbuk1, extra=8, uniform=FALSE, branch=1, left=FALSE, varlen=0, main="Min. 1 patient per end group")
rp.treegenclinbuk5 <- rpart(treegenclin$MRD~., method="class", data=treegenclin, control=rpart.control(minbucket=5, xval=10))
prp(rp.treegenclinbuk5, extra=8, uniform=FALSE, branch=1, left=FALSE, varlen=0, main="Min. 5 patient per end group")





#How many patients do we get right with our tree?
CorrectClass <- nrow(treegenclin) - MissClassErr*nrow(treegenclin) 
}
```
(2) Calculate Missclassification Error rate for those trees
    - How many patients do we get right with our tree?

(3) Relate bucket size to Missclassification Error

```{r}

Compare all four boosted models:
```{r}
source("multiplotfunction.R")
source("lastfunction.R")
multiplot(fancyRpartPlot(rp.treegenclin, uniform=FALSE, branch=1, left=FALSE))
