HICF1 - factor modelling - UNIVARIATES
========================================================

Dr. Susanne Weller 
14/05/2014

Import merged dataframe

```{r}
setwd("~/work/HICF1/HICF1_sub1/trunk/HICF_predictivemodel/HICF1_univariates")
load("genclin.Rda")

#convert integers to factors

genclin[,5] <- as.factor(genclin[,5])
genclin[,12:45] <- lapply(genclin[,12:45], as.factor)
```


Univariate analysis of variable effect on MRD


```{r}
#Chi-square test for factor variables

clones$Total.number.of.CNAs <- as.factor(clones$Total.number.of.CNAs)
Chi_CNAfreq <- table(clones$MRD, clones$Total.number.of.CNAs > 4)

cutoff <- as.list(c(4, 5, 6, 7, 8, 12))
output_CNAs <- vector("numeric", 61) 
for (i in 1:61){
Chi_CNAfreq <-table(clones$MRD, clones$Total.number.of.CNAs > i)
print(Chi_CNAfreq)
Xsq <- chisq.test(Chi_CNAfreq)
output_CNAs[i] <- Xsq[3]
}

#Wilcoxon test for continuous variables


# graph to show p-values for chi-square test against Total number of CNAs:
output_CNAs <- unlist(output_CNAs)
Total.number.of.CNAs <- c(1:61)
pvalues <- as.data.frame(cbind(output_CNAs, Total.number.of.CNAs))
pvalues <- subset(pvalues, Total.number.of.CNAs < 27)
```{r}
clones$TP53del <- factor(clones$TP53del, labels = c("TP53_not_del", "TP53_del"))
clones$TP53mut <- factor(clones$TP53mut, labels = c("TP53_not_mut", "TP53_mut"))

```


clones <-na.omit(clones[,c("MRD","Total.number.of.CNAs")])
m <- ggplot(clones, aes(clones$MRD, clones$Total.number.of.CNAs))
give.n <- function(x){
   return(c(y = mean(x), label = length(x)))
}
m + geom_boxplot(col=c("black", "purple"))+
  stat_summary(fun.data = give.n, geom = "text")+
  ggtitle("MRD ~ CNAs")+
  xlab("MRD")+
  ylab("Total number of CNAs")

#Violine plot so show outliers better:

r <- ggplot(clones, aes(clones$MRD, clones$Total.number.of.CNAs))
give.n <- function(x){
   return(c(y = mean(x), label = length(x)))
}
r + geom_violin(fill = "grey80", colour = "darkblue")+
  stat_summary(fun.data = give.n, geom = "text")+
  ggtitle("MRD ~ CNAs")+
  xlab("MRD")+
  ylab("Total number of CNAs")
pairwise.wilcox.test(clones$Total.number.of.CNAs, clones$MRD, p.adjust.method="bonferroni", paired = FALSE)
```

Subclones and MRD
```{r}

clones <-na.omit(clones[,c("MRD","Subclones")])
m <- ggplot(clones, aes(clones$MRD, clones$Subclones))
give.n <- function(x){
   return(c(y = mean(x), label = length(x)))
}
m + geom_boxplot(col=c("black", "purple"))+
  stat_summary(fun.data = give.n, geom = "text")+
  ggtitle("MRD ~ CNAs")+
  xlab("MRD")+
  ylab("Total number of CNAs")

# Violine plot
r <- ggplot(clones, aes(clones$MRD, clones$Subclones))
give.n <- function(x){
   return(c(y = mean(x), label = length(x)))
}
r + geom_violin(fill = "grey80", colour = "#3366FF")+
  stat_summary(fun.data = give.n, geom = "text")+
  ggtitle("Number of subclones influences MRD outcome")+
  xlab("MRD")+
  ylab("Subclones")
pairwise.wilcox.test(clones$Subclones, clones$MRD, p.adjust.method="bonferroni", paired = FALSE)

```

Frequency analysis - TP53 and MRD
```{r}
clones$TP53del <- factor(clones$TP53del, labels = c("TP53_not_del", "TP53_del"))
clones$TP53mut <- factor(clones$TP53mut, labels = c("TP53_not_mut", "TP53_mut"))

sumclones_del <-as.data.frame(table(clones$TP53del,clones$MRD))
sumclones_mut <-as.data.frame(table(clones$TP53mut, clones$MRD))

mut_M <- as.table(rbind(c(110, 0), c(92, 15)))
dimnames(mut_M) <- list(MRD = c("MRD_negativ","MRD_positive"),
                    TP53_status = c("TP53_not_mut","TP53_mut"))
(Xsq <- chisq.test(mut_M))  # Prints test summary
Xsq$observed   # observed counts (same as M)
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residuals

mosaicplot(mut_M)
library(vcd)
mosaic(mut_M, shade=TRUE, legend=FALSE, sub="Chi-Square test, p<0.0001***")
```

TP53 deletion and subclones

```{r}
### get p value
pvaldel <- wilcox.test(Subclones ~TP53del, data = clones)$p.value
### strip to 3 significant digits
pvaldel <- signif(pvaldel, 3)
### add asterisk if <0.05
pvaldel <- ifelse(pvaldel<0.06, paste0(pvaldel, " trend"), pvaldel)
pvaldel <- paste0("Wilcoxon Test \n p=", pvaldel)

r <- ggplot(clones, aes(clones$TP53del, clones$Subclones))
give.n <- function(x){
   return(c(y = mean(x), label = length(x)))
}
r + geom_violin(fill = "grey80", colour = "#3366FF")+
  stat_summary(fun.data = give.n, geom = "text")+
  ggtitle("TP53del and number of subclones")+
  xlab("TP53 deletion")+
  ylab("Subclones")

m <- ggplot(clones, aes(clones$TP53del, clones$Subclones))
give.n <- function(x){
   return(c(y = mean(x), label = length(x)))
}
m + geom_boxplot(col=c("black", "purple"))+
  stat_summary(fun.data = give.n, geom = "text")+
 ggtitle("TP53del and number of subclones")+
  xlab("TP53 deletion")+
  ylab("Subclones")+
  annotate("text",
             x=0.8,
             y=4,
             label=pvaldel)
pairwise.wilcox.test(clones$Subclones, clones$TP53del, p.adjust.method="bonferroni", paired = FALSE)
```
TP53 mutation and subclones
```{r}
r <- ggplot(clones, aes(clones$TP53mut, clones$Subclones))
give.n <- function(x){
   return(c(y = mean(x), label = length(x)))
}
r + geom_violin(fill = "grey80", colour = "#3366FF")+
  stat_summary(fun.data = give.n, geom = "text")+
  ggtitle("TP53 mutation and number of subclones")+
  xlab("TP53 mutation")+
  ylab("Subclones")+

### get p value
pvalmut <- wilcox.test(Subclones ~TP53mut, data = clones)$p.value
### strip to 3 significant digits
pvalmut <- signif(pvalmut, 3)
### add asterisk if <0.05
pvalmut <- ifelse(pvalmut<0.05, paste0(pvalmut, " *"), pvalmut)
pvalmut <- paste0("Wilcoxon Test \n p=", pvalmut)

### plot as per OP
m <- ggplot(clones, aes(clones$TP53mut, clones$Subclones))
give.n <- function(x){
   return(c(y = mean(x), label = length(x)))
}
m + geom_boxplot(col=c("black", "purple"))+
  stat_summary(fun.data = give.n, geom = "text")+
 ggtitle("TP53 mutation and number of subclones")+
  xlab("TP53 mutation")+
  ylab("Subclones")+
    annotate("text",
             x=0.8,
             y=4,
             label=pvalmut)


```
Frequencies of "Total number of CNAs" who are "MRD" positive or negative

```{r}

CNAfreq <-as.data.frame(table(clones$MRD,clones$Total.number.of.CNAs))
colnames(CNAfreq) <- c("MRD", "Total.number.of.CNAs", "Frequency")

# Stacked bar chart using ggplot
clones <-na.omit(clones[,c("MRD","Subclones", "Total.number.of.CNAs")])
k <- ggplot(clones, aes(factor(clones$Total.number.of.CNAs), fill=clones$MRD))
k + geom_bar()+
ggtitle("Number of CNAs within CLL sample set")+
  xlab("Total number of CNAs")+
  ylab("Number of samples")

#dodged bar chart
n <- ggplot(clones, aes(factor(clones$Total.number.of.CNAs),fill=clones$MRD))
n+geom_bar(position="dodge")+
 ggtitle("Number of CNAs within CLL sample set")+
  xlab("Total number of CNAs")+
  ylab("Number of samples")+
  scale_fill_manual(values=c("darkblue", "darkred"))

# Chi - square test with different cut-offs for what CNAs are considered as "normal"

clones$Total.number.of.CNAs <- as.factor(clones$Total.number.of.CNAs)
Chi_CNAfreq <- table(clones$MRD, clones$Total.number.of.CNAs > 4)

cutoff <- as.list(c(4, 5, 6, 7, 8, 12))
output_CNAs <- vector("numeric", 61) 
for (i in 1:61){
Chi_CNAfreq <-table(clones$MRD, clones$Total.number.of.CNAs > i)
print(Chi_CNAfreq)
Xsq <- chisq.test(Chi_CNAfreq)
output_CNAs[i] <- Xsq[3]
}

# graph to show p-values for chi-square test against Total number of CNAs:
output_CNAs <- unlist(output_CNAs)
Total.number.of.CNAs <- c(1:61)
pvalues <- as.data.frame(cbind(output_CNAs, Total.number.of.CNAs))
pvalues <- subset(pvalues, Total.number.of.CNAs < 27)

p <- ggplot(pvalues, aes(x=pvalues$Total.number.of.CNAs, y=pvalues$output_CNAs))
p+geom_point(shape=1)+
  scale_y_reverse()+
  geom_hline(yintercept = 0.05)+
  scale_x_continuous(breaks=seq(1, 45, 1))+
  ggtitle("Significance of CNAs cutoffs")+
  xlab("Total number of CNAs - cutoff")+
  ylab("p-value for Chi-square test") 


# Mosaic plot for Cutoff at 6

Chi_CNAfreq_6 <- table(clones$MRD, clones$Total.number.of.CNAs > 6)
dimnames(Chi_CNAfreq_6) <- list(MRD = c("MRD_negativ","MRD_positive"),
                    Total_number_of_CNAs = c(">/= 6","< 6"))

mosaicplot(Chi_CNAfreq_6, shade=TRUE, legend=FALSE, sub="Chi-Square test, p=0.047")
library(vcd)
mosaic(Chi_CNAfreq_6, shade=TRUE, legend=FALSE, sub="Chi-Square test, p=0.047")

